Deployable Probabilistic Programming

### 2019-04-18 paper/paper.tex
inference drafted

@@

Any user-defined function of appropriate kind can be used as an
elemental inside model methods. An elemental must be declared
for the function using a call to
\begin{lstlisting}
func RegisterElemental(
	f interface{}, 
	g func(value float64,
		   params ...float64) []float64).
\end{lstlisting} For
example, the call
\begin{lstlisting}
RegisterElemental(Sigm,
	func(value float64,
	     _ ...float64) []float64 {
		return []float64{value*(1-value)}
	})
\end{lstlisting}
registers the gradient for function \lstinline{Sigm}.

@@

When only a point estimate of the model parameters is required,
as, for example, in the simulation-inference case, stochastic
gradient descent is a fast and robust inference family. Infergo
offers stochastic gradient descent with momentum and
Adam~\cite{KB15}.  The inference is performed by calling the
\lstinline{Step} method of the optimizer until a termination
condition is met.
\begin{lstlisting}
\end{lstlisting}

When the full posterior must be recovered for integration,
Hamiltonian Monte Carlo (HMC) can be used instead. Vanilla HMC
as well as the No-U-Turn sampler (NUTS)~\cite{HG11} are
provided. To support lazy any-time computation, the inference
runs in a separate \textit{goroutine}, connected by a channel to
the caller, and asynchronously writes samples to the channel.
The caller reads as many samples from the channel as needed, and
further post-processes them.
\begin{lstlisting}
\end{lstlisting}

Other inference schemes, most notably automatic differentiation
variational inference~\cite{KTR+17}, are planned for addition
in future versions of Infergo. However, an inference algorithm
does not have to be a part of Infergo to work with Infergo
models. Since infergo models operate on the built-in Go floating
point type, \lstinline{float64}, any third-party inference or
optimization library can be used as well.

As an example, the Gonum~\cite{Gonum} library for numerical
computations contains an optimization package offering several
algorithms for gradient-based optimization. A Gonum optimization
algorithm require the function to optimize as well as the
function's gradient as parameters. An Infergo model can be
trivially wrapped for use with Gonum, and Infergo provides a
convenience function
\begin{lstlisting}
func FuncGrad(m Model) (
	Func func(x []float64) float64,
	Grad func(grad []float64, x []float64)
		[]float64)
\end{lstlisting}
which accepts a model and returns the model's
\lstinline{Observe} method and its gradient suitable for passing
to Gonum optimization algorithms. Among available algorithms are
advanced versions of stochastic gradient descent, as well as
algorithms of BFGS family, in particular LBFGS~\cite{LN89},
providing for a more efficient MAP estimation than stochastic
gradient descent at the cost of higher memory consumption.

@@

>>>>

%\linespread{1.0}

@@

Any user-defined function of appropriate kind can be used as an
elemental inside model methods. An elemental must be declared
for the function using a call to
\begin{lstlisting}
func RegisterElemental(
	f interface{}, 
	g func(value float64,
		   params ...float64) []float64)
\end{lstlisting} For
example, the call
\begin{lstlisting}
RegisterElemental(Sigm,
	func(value float64,
	     _ ...float64) []float64 {
		return []float64{value*(1-value)}
	})
\end{lstlisting}
registers the gradient for function \lstinline{Sigm}.

@@

When only a point estimate of the model parameters is required,
as, for example, in the simulation-inference case, stochastic
gradient descent is a fast and robust inference family. Infergo
offers stochastic gradient descent with momentum and
Adam~\cite{KB15}.  The inference is performed by calling the
\lstinline{Step} method of the optimizer until a termination
condition is met:
\begin{lstlisting}
for !<termination condition> {
    optimizer.Step(model, parameters)
}
\end{lstlisting}

When the full posterior must be recovered for integration,
Hamiltonian Monte Carlo (HMC) can be used instead. Vanilla HMC
as well as the No-U-Turn sampler (NUTS)~\cite{HG11} are
provided. To support lazy any-time computation, the inference
runs in a separate \textit{goroutine}, connected by a channel to
the caller, and asynchronously writes samples to the channel.
The caller reads as many samples from the channel as needed, and
further postprocesses them:
\begin{lstlisting}
samples := make(chan []float64)
hmc.Sample(model, parameters, samples)
for !<termination condition> {
	sample = <-samples
	// postprocess the sample
}
hmc.Stop()
\end{lstlisting}

Other inference schemes, most notably automatic differentiation
variational inference~\cite{KTR+17}, are planned for addition
in future versions of Infergo. However, an inference algorithm
does not have to be a part of Infergo to work with Infergo
models. Since Infergo models operate on a built-in Go floating
point type, \lstinline{float64}, any third-party inference or
optimization library can be used as well.

As an example, the Gonum~\cite{Gonum} library for numerical
computations contains an optimization package offering several
algorithms for gradient-based optimization. A Gonum optimization
algorithm requires the function to optimize, as well as the
function's gradient, as parameters. An Infergo model can be
trivially wrapped for use with Gonum, and Infergo provides a
convenience function
\begin{lstlisting}
func FuncGrad(m Model) (
	Func func(x []float64) float64,
	Grad func(grad []float64, x []float64)
		[]float64)
\end{lstlisting}
which accepts a model and returns the model's
\lstinline{Observe} method and its gradient suitable for passing
to Gonum optimization algorithms. Among available algorithms are
advanced versions of stochastic gradient descent, as well as
algorithms of BFGS family, in particular LBFGS~\cite{LN89},
providing for a more efficient MAP estimation than stochastic
gradient descent at the cost of higher memory consumption:
\begin{lstlisting}
function, gradient := FuncGrad(model)
problem := gonum.Problem{Func: function, 
                         Grad: gradient}
result, err := gonum.Minimize(problem, parameters)
\end{lstlisting}

@@

\clearpage
\clearpage
